---
title: "Nobel Prize API"
author: "Guibril Ramde"
format: html
editor: visual
---

## ##Approach

In this assignment, I will use the Nobel Prize API to retrieve JSON data about Nobel laureates and prizes. The main endpoint I will use is the Laureates API, which returns information about individuals and organizations connected to Nobel Prizes. I may also use the Nobel Prizes endpoint if needed for prize-level comparisons. The Nobel Prize Developer Zone provides both of these public endpoints in API version 2.1\
\
After retrieving the JSON data in R, I will transform the nested response into tidy data frames using packages such as \`jsonlite\`, \`httr\`, \`dplyr\`, \`tidyr\`, and \`janitor\`. I will then explore the cleaned data to answer four data-driven questions about Nobel laureates.\
\
The four questions I plan to investigate are:

-   Which countries have produced the highest number of Nobel laureates?

-   Which country has lost the most Nobel laureates, meaning laureates who were born there but received the prize as citizens of another country?

-   Have any individuals won the Nobel Prize more than once?

-   What age group is most common among Nobel laureates at the time they received the prize?

    For each question, I will describe the question, show the R code used to answer it, and present the result using a table, summary, or visualization.

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(jsonlite)
library(httr)
library(janitor)


url <- "https://api.nobelprize.org/2.1/laureates"

response <- GET(url)
stop_for_status(response)

raw_text <- content(response, as = "text", encoding = "UTF-8")
data <- fromJSON(raw_text, flatten = TRUE)


names(data)
glimpse(data)


```

```{r}
laureates <- as_tibble(data$laureates) |>
  clean_names()

glimpse(laureates)
```
