---
title: "Sentiment Analysis"
author: "Guibril Ramde"
format: html
editor: visual
---

## ##Approach

In this assignment, I will reproduce and extend the sentiment analysis example from Chapter 2 of \*Text Mining with R\* by Julia Silge and David Robinson. First, I will recreate the original example in a Quarto document using the same general workflow shown in the chapter. This includes tokenizing the text, joining it with sentiment lexicons, and analyzing the emotional tone of the text.\
\
To reproduce the base example, I will use the \`janeaustenr\` package together with the \`tidytext\` and \`textdata\` packages. I will download the required sentiment lexicons such as AFINN, Bing, and NRC, and then use them to analyze the Jane Austen texts.\
\
To extend the analysis, I will use a different text corpus of my choice and apply at least one additional sentiment lexicon. I will then compare the results from the new corpus with the original Jane Austen example and explain how the sentiment patterns differ across lexicons and texts.\
\
For the final report, I will include:\
1. the reproduced Chapter 2 example,\
2. a citation to the book and original example,\
3. an extended analysis using a new corpus,\
4. at least one additional sentiment lexicon,\
5. and an explanation of how the new results differ from the original results.

```{r}
install.packages("textdata")
library(tidytext)
library(textdata)

get_sentiments("afinn")
```

```{r}
get_sentiments("bing")
```

```{r}
get_sentiments("nrc")
```

```{r}
library(janeaustenr)
library(stringr)


tidy_books <- austen_books() |>
  group_by(book) |>
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text,
                                regex("^chapter [\\divxlc]",
                                            ignore_case = TRUE)))) |>
  ungroup() |>
  unnest_tokens(word, text)

nrc_joy <- get_sentiments("nrc") |>
  filter(sentiment == "joy")


tidy_books |>
  filter(book == "Emma") |>
  inner_join(nrc_joy) |>
  count(word, sort = TRUE)
  
```
